# Missing values

Below are some key observations with respect to Missing Values in the dataset. 

```{r}
library(tidyverse)
library(patchwork)
library(ggplot2)
library(mi)
library(extracat)
library(dplyr)
```

```{r}
theme_fakeNews_rotate <- function(base_size = 14, ang=90) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = 25, face = "bold", margin = margin(10,10,10,10), hjust = 0.5),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.x = element_text(size = 13, face = "bold", margin = margin(10,10,20,10), angle = ang),
      axis.text.y = element_text(size = 13, face="bold", margin = margin(10,10,10,10)),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # La légende
      legend.title = element_text(size = 13, face = "bold"),
      legend.text = element_text(size = 13, face = "bold")
    )}

theme_fakeNews <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = 25, face = "bold", margin = margin(10,10,10,10), hjust = 0.5),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.x = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.y = element_text(size = 13, face="bold", margin = margin(10,10,10,10)),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # La légende
      legend.title = element_text(size = 13, face = "bold"),
      legend.text = element_text(size = 13, face = "bold")
    )}
```

```{r}
missingGraphPropotion <- function(df, percent=FALSE, shorten=FALSE) {
  
  # initializing variables depending on percentage or count
  alph <- 0.4
  p1_left <- 'num'
  p2_bot <- 'row count'
  limit <- NULL
  
  if(percent){
    p1_left <- '%'
    p2_bot <- '% rows'
    limit <- c(0,100)
  }
  sh <- waiver()
  if(shorten){
    sh <- abbreviate
  }

  # gets null proportion per column
  hor.index <- function(df) {
    v <- c()
    for (i in head(names(df), -1)) {
      if(percent){
        frac <- 100*sum(df$count[c(df[,i]==TRUE)])/sum(df$count)
      } else {
        frac <- sum(df$count[c(df[,i]==TRUE)])
      }
      v <- c(v, frac)
    }
    data <- data.frame(
      name = head(names(df), -1),
      value = v
    )
    return(data)
  }
  
  # organizes dataframe by null pattern, and uses the null porportion to reorder columns
  null.graph <- function(df) {
    ms <- data.frame(is.na(df)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup()
    
    ct <- ms$count
    ind <- hor.index(ms)$value
    ms <- ms[, order(ind, decreasing = TRUE)]
    ms$count = ct
    return(ms)
  }
  
  g <- null.graph(df)
  ind <- hor.index(g)
  
  
  # top side plot
  p1 <- ggplot(data = ind, aes(x = reorder(name, desc(value)), y = value)) + 
    geom_bar(color='white', stat = 'identity', alpha = alph) +
    ggtitle('Missing value patterns') +
    ylab(paste(p1_left, 'rows missing:')) +
    xlab('') +
    scale_y_continuous(expand = c(0,0), limits = limit, n.breaks = 5) +
    scale_x_discrete(labels = sh) +
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = 25, face = "bold", margin = margin(10,10,10,10), hjust = 0.5),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.x = element_text(size = 13, face = "bold", margin = margin(10,10,10,10), angle = 90),
      axis.text.y = element_text(size = 13, face="bold", margin = margin(10,10,10,10)),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # La légende
      legend.title = element_text(size = 13, face = "bold"),
      legend.text = element_text(size = 13, face = "bold"))
  
  if(percent){
    g$count <- 100*g$count/sum(g$count)
  }

  
  # make pl dataframe for the main plot
  g$row=as.numeric(row.names(g))
  pl <- pivot_longer(data = g, 
               cols = -c(count, row),
               names_to = "name",
               values_to = "vals")
  pl <- merge(pl, ind)
  
  
  # make fill values for alpha
  g2 <- g[1:(length(g)-2)]
  empty <- c()
  for (row in 1:nrow(g2)) {
    empty <- c(empty, sum(g2[row,]))
  }
  fill <- data.frame(
    row = c(1:nrow(g2)),
    fill = empty
  )
  
  
  # merge fill values to pl, and add fill column to g
  pl <- merge(pl, fill)
  g$fill <- fill$fill
  
  # if there are no missing columns
  if(sum(g$fill)==0){
    alph = 1
  }
  
  # right side plot
  p2 <- ggplot(data = g, aes(x = reorder(as.numeric(row.names(g)), desc(as.numeric(row.names(g)))), y = count, alpha = fill == 0)) + 
    geom_bar(color='white', stat = 'identity') +
    ylab(p2_bot) +
    xlab('') +
    scale_y_continuous(expand = c(0,0), limits = limit, n.breaks = 5) +
    scale_alpha_manual(values = c(alph, 1)) +
    coord_flip() +
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = 25, face = "bold", margin = margin(10,10,10,10), hjust = 0.5),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.x = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.y = element_text(size = 13, face="bold", margin = margin(10,10,10,10)),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # La légende
      legend.title = element_text(size = 13, face = "bold"),
      legend.text = element_text(size = 13, face = "bold"))
  
  
  # main plot
  p3 <- ggplot(data = pl, aes(x = reorder(name, desc(value)), y = reorder(row, desc(row)), fill = vals, alpha = fill == 0)) +
  geom_tile(color = 'white') +
  scale_alpha_manual(values = c(alph, 1)) +
  annotate('text', x = length(x = unique(pl$name))/2 + 0.5, y = tail(pl$row,1) - pl[pl$fill == 0,]$row[1] + 1, label = 'Complete Cases') +
  scale_x_discrete(labels = sh) +
  theme(
      # L'ensemble de la figure
      plot.title = element_text(size = 25, face = "bold", margin = margin(10,10,10,10), hjust = 0.5),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = 13, face = "bold", margin = margin(10,10,10,10)),
      axis.text.x = element_text(size = 13, face = "bold", margin = margin(10,10,10,10), angle = 90),
      axis.text.y = element_text(size = 13, face="bold", margin = margin(10,10,10,10)),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # La légende
      legend.title = element_text(size = 13, face = "bold"),
      legend.text = element_text(size = 13, face = "bold")) +
  xlab('variable') +
  ylab('missing pattern')
  
  
  # combine plots with patchwork
  pp <- p1 + plot_spacer() + p3 + p2 +
    plot_layout(widths = c(5,1), heights = c(1,4))
  
  return(pp)
}
```



```{r}
df_missing = read_csv(file = 'FakeNews_Unclean.csv' )
```



**Missing values by column for the Liar dataset**

```{r}
colSums(is.na(df_missing)) %>%
  sort(decreasing = TRUE)
```


1) It is important to note that Label column (independent feature) doesn't have missing values in the dataset. 

2) Most of the columns have 2 missing values, which after taking a closer look belongs to ID's 5872 and 8180

3) Speaker Job Title has maximum missing values - 3170


**Top 30 row-id having maximum missing values**

```{r}
temp_df = df_missing
row.names(temp_df) <- 1 : 11378
t = rowSums(is.na(temp_df)) %>%
  sort(decreasing = TRUE)

head(t,30)
```


1) Maximum number of columns having missing values in a row are 11. For examples Id's like 5872, 8180 etc. have 11 columns with missing values.


```{r echo=FALSE, message=FALSE, results='hide'} 

#Ratio of Missing Values for each Label in Test dataset 

features = c("Label", "Speaker Job Title")
df_new = df_missing[features]

df_new  = df_missing %>%
group_by(Label) %>%
summarise_each(funs(sum(is.na(.))/length(.)))

df_new1 = df_new[order(df_new$`Speaker Job Title`, decreasing=TRUE),]

barplot(df_new1$`Speaker Job Title`*100 , 
        names.arg = df_new1$Label,
        xlab = "Label of fake news",
        ylab = "Percentage of Missing Values",
        main ="Percentage of missing values with respect to Label") + theme_fakeNews()

```


**Heatmap **

```{r fig.width=10}
df_pivot <- df_missing %>%
    gather(key, value, -Label) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(df_pivot, aes(y = key, x =  fct_rev(Label), fill = missing)) +
  geom_tile(color = "white") + 
  labs(title = "Missing Values for Features Faceted by Label",
       x = "Label Associated", y="Features in the Dataset") + theme_fakeNews()
```


Since we noticed that "Speaker Job Title" and "State Info" had maximum missing values, we used heatmaps to see if we can observe any pattern.     Below are the observed patterns :-

  1) For "Speaker Job Title", True & False Staements  were completely missing.
  
  2) For "State Info" feature, False Statements were completely missing. 



**Missing propotions**

```{r fig.width=17, fig.height=10}

missingGraphPropotion(df_missing) 
```

